\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-06-06 final-version/concepts.tex
updates to final version

\paragraph {A brief note on \eu history and notation} The origin of
what we will call \eu in this chapter, is derived from the
\texttt{HSTS} planner \cite{mus94} used on a number of ground-based
demonstrations for space telescope scheduling as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was also the \texttt{LISP}-based
planner that was flown as part of the Remote Agent \texttt{RAX}
experiment. Concurrent to \texttt{RAX} deployment, a C++ version of
the planner was implemented \kcomment{at NASA Ames. This} went on to
be used for the \texttt{MAPGEN} \cite{bresina05} system and continues
to be used to date, in the mission-critical uplink process for the
command/control of the Mars Exploration Rovers (\texttt{MER})
mission. A vastly re-factored, higher performance open source version
\cite{europapso} was named \eut and is the subject of this chapter and
used by \rxe.  Note however, in using the name \eue, we refer to the
\emph{fundamental ideas as well as the planning paradigm} even if
there exists a specific implementation.

>>>>

\paragraph {A brief note on \eu history and notation} The origin of
what we will call \eu in this chapter, is derived from the
\texttt{HSTS} planner \cite{mus94} used on a number of ground-based
demonstrations for space telescope scheduling as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was also the \texttt{LISP}-based
planner that was flown as part of the Remote Agent \texttt{RAX}
experiment \cite{mus98}. Concurrent to the \texttt{RAX} deployment, a
C++ version of the planner was implemented \kcomment{at NASA
  Ames. This} went on to be used for the \texttt{MAPGEN} system and
continues to be used to date, in the mission-critical uplink process
for the command/control of the Mars Exploration Rovers (\texttt{MER})
mission \cite{bresina05}. A vastly re-factored, higher performance
open source version was named \eut \cite{europapso} and is the subject
of this chapter and used by \rxe.  Note however, in using the name
\eue, we refer to the \emph{fundamental ideas as well as the planning
  paradigm}.

### 2012-06-06 final-version/intro.tex
updates to final version

@@

In this chapter, we propose an alternative view to decision-making
using in-situ automated Planning \cite{ghallab04}. More importantly,
we demonstrate that earlier control paradigms (primarily driven by
\emph{reactive} approaches to control) can be augmented to provide
targeted observation and sampling as well as a more nuanced and
balanced consideration of mission objectives, environmental conditions
and available resources. Planning (or deliberation or projection in
action space\footnote{\kcomment{We use the terms 'planning' and
    'deliberation' interchangeably in this chapter.}}) is important to
balance current needs of a robot with future desires or goals taking
into account time and available resources in-situ on the
robot. Planning without time or action planning in and of itself is
not adequate; dealing with time ensures that the balance between now
and the future is handled systematically. Yet, plans have to be tied
to robotic action and in dynamic real-world
\kcomment{environments}. Further, plans \kcomment{not only} project
future state which are likely to change, but \kcomment{need to factor
  in} substantial environmental uncertainty\footnote{It is important
  to remember that these principles are broad much as Dwight
  Eisenhower is reputed to have said ``\emph{In preparing for battle I
    have always found that plans are useless, but planning is
    indispensable}'' and ``\emph{Failing to plan is planning to
    fail}''.}. \kcomment{In the oceanographic context,} use of
environmental cues is often hampered by poor \kcomment{predictive}
skill \cite{anderson2009} of ocean models and little to no
availability of a priori synoptic views of the survey
area. \kcomment{Planning therefore provides a ``buffer'' against the
  stochastic environment.}

@@

This chapter is organized as follows. We start \kcomment{by placing
  this effort in the context of the overal research in AI Planning in
  Section \ref{sec:related}} followed by some foundational concepts in
our Planning framework in Section~\ref{sec:concepts} which includes a
brief overview of AI Planning in Section~\ref{sec:planningfound}. In
Section~\ref{sec:basics} we go deeper into the \eu framework itself,
the basis of all deliberation within \rxe. A key portion is
Section~\ref{sec:europa:pr} which details the internal plan
representation within a situated agent like \rxe. Modeling
(Section~\ref{sec:europa:modeling}), Inference
(Section~\ref{sec:europa:inference}) and Search
(Section~\ref{sec:europa:search}) articulate important details
associated with \eu critical to \rxe.  With these foundational
elements of deliberation out of the way, we transition to \rx itself
in Section~\ref{sec:arch}, bring out architectural details in
Section~\ref{sec:arch:trex}, deal with the execution cycle in
Section~\ref{sec:arch:exec}, highlight how \rx deliberates in
Section~\ref{sec:arch:europa}, finally delving deeper into the key
concept of synchronization in Section~\ref{sec:arch:synch}. Further
details on deliberation within \rx (Section~\ref{sec:arch:plan}) and
how planning and execution are interleaved are detailed in
Section~\ref{sec:arch:intertwine}. We offer results from our
experiments at sea in Section~\ref{sec:results}, offer some insights
into the near-term future goals of our research in
Section~\ref{sec:future} and conclude with
Section~\ref{sec:conclusion}.

>>>>

\setcounter{footnote}{0}

@@

In this chapter, we propose an alternative view to decision-making
using in-situ automated Planning \cite{ghallab04}. More importantly,
we demonstrate that earlier control paradigms (primarily driven by
\emph{reactive} approaches to control) can be augmented to provide
targeted observation and sampling as well as a more nuanced and
balanced consideration of mission objectives, environmental conditions
and available resources. Planning\footnote{We use the terms 'planning'
  and 'deliberation' interchangeably in this chapter.}  (or
deliberation or projection in action space) is important to balance
current needs of a robot with future desires or goals taking into
account time and available resources in-situ on the robot. Planning
without time or action planning in and of itself is not adequate;
dealing with time ensures that the balance between now and the future
is handled systematically. Yet, plans have to be tied to robotic
action and in dynamic real-world \kcomment{environments}. Further,
plans \kcomment{not only} project future state which are likely to
change, but \kcomment{need to factor in} substantial environmental
uncertainty\footnote{It is important to remember that these principles
  are broad much as Dwight Eisenhower is reputed to have said
  ``\emph{In preparing for battle I have always found that plans are
    useless, but planning is indispensable}'' and ``\emph{Failing to
    plan is planning to fail}''.}. \kcomment{In the oceanographic
  context,} use of environmental cues is often hampered by poor
\kcomment{predictive} skill \cite{anderson2009} of ocean models and
little to no availability of a priori synoptic views of the survey
area. \kcomment{Planning therefore provides a ``buffer'' against the
  stochastic environment.}

@@

This chapter is organized as follows. We start \kcomment{by placing
  this effort in the context of the overall research in AI Planning in
  Section \ref{sec:related}} followed by some foundational concepts in
our planning framework in Section~\ref{sec:concepts} which includes a
brief overview of AI Planning in Section~\ref{sec:planningfound}. In
Section~\ref{sec:basics} we go deeper into the \eu framework itself,
the basis of all deliberation within \rxe. A key portion is
Section~\ref{sec:europa:pr} which details the internal plan
representation within a situated agent like \rxe. Modeling
(Section~\ref{sec:europa:modeling}), Inference
(Section~\ref{sec:europa:inference}) and Search
(Section~\ref{sec:europa:search}) articulate important details
associated with \eu critical to \rxe.  With these foundational
elements of deliberation out of the way, we transition to \rx itself
in Section~\ref{sec:arch}, bring out architectural details in
Section~\ref{sec:arch:trex}, deal with the execution cycle in
Section~\ref{sec:arch:exec}, highlight how \rx deliberates in
Section~\ref{sec:arch:europa}, finally delving deeper into the key
concept of synchronization in Section~\ref{sec:arch:synch}. Further
details on deliberation within \rx (Section~\ref{sec:arch:plan}) and
how planning and execution are interleaved are detailed in
Section~\ref{sec:arch:intertwine}. We offer results from our
experiments at sea in Section~\ref{sec:results}, some insights into
the near-term future goals of our research in Section~\ref{sec:future}
and conclude with Section~\ref{sec:conclusion}.

### 2012-06-06 final-version/related.tex
updates to final version

A distinct problem in Marine Robotics has been the use of the
overloaded term 'autonomy'. Not only does the notion transcend
different disciplines within engineering in this domain (e.g. the
Control Engineering community sees it distinctively from those in AI)
but users of marine platforms in oceanography as well as in maritime
defense conflate the methodology of use (tethered vs. untethered) with
their control. In this chapter, our notion of 'autonomy' deals
explicitly not only with the notion of dealing with the \emph{sense,
  plan and act} paradigm, but also the need to perform
\emph{computational search} between different action outcomes, an idea
central to the field of AI. Consequently, our literature survey here
is targeted towards a more focused and (to this chapter) relevant part
of the field of computation.

>>>>

A distinct problem in Marine Robotics has been the use of the
overloaded term 'autonomy'. Not only does the notion transcend
different disciplines within engineering in this domain (e.g. the
Control Engineering community sees it distinctively from those in AI)
but users of marine platforms in oceanography as well as in maritime
defense conflate the methodology of use (tethered vs. untethered) with
their control. In this chapter, our notion of 'autonomy' refers not
only to the notion of dealing with the \emph{sense, plan and act}
paradigm in robotics, but also the need to perform \emph{computational
  search} between different action outcomes, an idea central to the
field of AI. Consequently, our literature survey here is targeted
towards a more focused and (to this chapter) relevant part of the
field of computation.
